\LARGE \bf ***

### 2020-09-24 icra-2021.tex
fixed up until ...computations (proofreading now)

The proposed algorithm alters the energy consumption dynamically by means of mission-specific parameters: the Quality of Service (QoS) of the onboard computations, and the trajectory-explicit equations (TEEs) of the path. Our goal is a mission extension by optimizing both QoS and TEEs as the UAV flies and its batteries drain. The algorithm optimizes QoS requiring the UAV to include robot operating system (ROS) nodes. It optimizes TEEs--an abstraction of the path to follow--and moves the UAV using a vector field that converges smoothly to the path. Physically, one can use a feature detection node and a set of ellipses, with the algorithm controlling the processing rate and the length of the semi-major and -minor axis. Data using a commercial fixed-wing UAV indicates a potential extension of up to 13 minutes over an hour by merely switching to the lowest QoS.

@@

The remainder of the paper is organized as follows. The dynamic mission planning problem is set in Section~\ref{sec:prob}. A suitable model for the energy and position of the UAV is showed in Section~\ref{sec:model}. The algorithm that uses the model and solves the problem is proposed in Section~\ref{sec:algo}. Section~\ref{sec:experimental} presents a fixed-wing UAV flying in an agricultural scenario showcasing the performance. The paper finishes with some conclusions in Section~\ref{sec:conclusion}.

@@

The UAV flies in ellipses shifted in time--a survey pattern optimized for the craft's dynamics--detects obstacles using a convolutional neural network (CNN) and informs ground mobile robots employed for future harvesting.

@@

We use the state to estimate the UAV's position in space and the energy evolution in time. Starting with a position $\mathbf{p}\in\mathbb{R}^3$ in the 3D Euclidean space w.r.t. some inertial navigation frame $\mathcal{O}_W$, we build a guidance action--which allows the \emph{position evolution} along the 3-axis--to the path to follow using a vector field~\cite{de2017guidance}. The path is a set of generic continuous twice differentiable TEEs $\varphi_k:\mathbb{R}^3\rightarrow\mathbb{R}$. To this end, the TEE is satisfied $\varphi(\mathbf{p})\rightarrow 0$ for all the approaching points.

We build the \emph{energy evolution} $\mathbf{q}\in\mathbb{R}^j$ using Fourier analysis (the meaning of $j$ is clarified to the reader in Subsection~\ref{sec:energy-model}) and decompose such evolution in the energy due to the trajectory (depending on TEEs), and the computations (on QoS)--an approach adapted from our earlier work on computational energy analysis~\cite{seewald2019coarse, seewald2019component}, and energy estimation of a fixed-wing UAV~\cite{seewald2020mechanical}.

\subsection{Position in space and guidance action}
\label{sec:position}

We consider a non-holonomic 2D model of the UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ (the requirement is eased in Section~\ref{sec:algo} to a set of altitudes $\mathbf{h}$)
\begin{equation}\label{eq:uav-kinem}\begin{cases}
  \dot{\mathbf{p}}(t)&=s\Psi(\psi(t))+\mathbf{d}(t)\\
  \dot{\psi}(t)&=u(\mathbf{p}(t))
\end{cases},
\end{equation}
where $\mathbf{p}\in\mathbb{R}^2$, $s\in\mathbb{R}$ is the airspeed (assumed constant), $\psi\in(-\pi,\pi]$ the attitude yaw angle and $\Psi(\psi)=\begin{bmatrix}\cos{\psi} & \sin{\psi}\end{bmatrix}^T$, $\mathbf{d}\in\mathbb{R}^2$ is the wind vector (assumed $\|\mathbf{d}\|<s$, i.e., the constant airspeed is greater than the norm of the wind), and $u\in\mathbb{R}$ the guidance action--the angular velocity $\dot{\psi}$ of the UAV.

Let us define $\mathcal{P}_k$, the area that discloses all the possible deviations form the $k$-th TEE 
\begin{equation}\label{eq:area}
  \mathcal{P}_k:=\{\mathbf{p}:\underline{c}_k\leq\varphi_k(\mathbf{p})\leq\overline{c}_k\},
\end{equation}
where $\underline{c}_k\in\mathbb{R}_{\leq 0},\overline{c}_k\in\mathbb{R}_{\geq 0}$ are the constraints. $\underline{a},\overline{a}$ returns the upper and lower bound of $a$ from the user-defined mission specification, a lookup table.

The algorithm uses the concept to select $c$ with the highest energy value under the energy budget constraints. Assume at time $k$ the algorithm selects the TEE $\varphi_k$ with the highest energy value being $c_k$, we design a guidance action $u$ that follows $\varphi_k$ by minimizing the norm $\|\varphi_k(\mathbf{p})-c_k\|$.

Let us define $\varPhi:=\varphi_k(\mathbf{p})-c_k$. The direction to follow can be expressed as the desired velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is defined as the gradient of $\varphi$ at the point $\mathbf{p}$ (i.e., its vector field), $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence.

@@

@@

To evaluate the energy we formalize the following realistic assumption introduced in Section~\ref{sec:intro}.
\begin{assm}\label{assm:periodic}
The mission the mobile robots perform is periodic, in the sense that it presents repetitive patterns.
\end{assm}

Let us consider a Fourier series $f:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}_{\geq 0}$ of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$
\begin{equation}\label{eq:fourier}
  f(t)=\sum_{n=0}^{r}{a_n\cos{\frac{nt}{\xi}}+b_n\sin{\frac{nt}{\xi}}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ for $n\in\{0,\dotsc,r\}$ the Fourier series coefficients.

The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using an equivalent time-varying state-space model in the following form
\begin{equation}\label{eq:state-perf}\begin{cases}
  \dot{\mathbf{q}}(t)&=A\mathbf{q}(t)+B\mathbf{u}(t)\\%+w(t)\\
  y(t)&=C\mathbf{q}(t)%+v(t)
\end{cases},\end{equation}
where $y(t)\in\mathbb{R}_{\geq 0}$ is the instantaneous energy evolution of the system being controlled. We will see later in the Theorem~\ref{thm:state-vs-energy} that the instantaneous energy can be obtained from the model as a linear combination of the state. The control $\mathbf{u}$ along with the input matrix $B$ are defined later in Subsection~\ref{sec:control-action}, the state $\mathbf{q}$ mimics the original Fourier series coefficients, and
\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}(t)&=\left[\begin{array}{ccccccc}
    \alpha_0 & \vline & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{ccccc}
    1          & \vline & \mathbf{0} & \cdots & \mathbf{0} \\\hline
    \mathbf{0} & \vline & A_1        & \cdots & \mathbf{0} \\
    \vdots     & \vline & \vdots     & \ddots & \vdots     \\
    \mathbf{0} & \vline & \mathbf{0} & \cdots & A_r 
  \end{array}\right],\,A_n=\begin{bmatrix}0 & \frac{n}{\xi} \\ -\frac{n^2}{\xi^2} & 0\end{bmatrix},\\
  C&=\left[\begin{array}{ccccccc}
    1 & \vline & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}(t)\in\mathbb{R}^j$ given $j:=2r+1$, $A\in\mathbb{R}^{j\times j}$ is the state transmission matrix, and $C\in\mathbb{R}^j$ is the output matrix. Furthermore, the first row and column of the matrix $A$ contain zeros row vectors and column vectors respectively.

Motivated by the fact that the system of interest is sampled discrete-time, we consider the discretized version of Equation~(\ref{eq:state-perf}) and add the uncertainty
\begin{equation}\label{eq:state-unct}
  \begin{cases}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+w_k\\
  y_k&=C\mathbf{q}_k+v_k
  \end{cases},
\end{equation}
where $w_k\in\mathbb{R}$ accounts for the environment uncertainty, and $v_k\in\mathbb{R}$ for the measurement error.

The instantaneous energy can be obtained from the model~(\ref{eq:state-unct}) as a linear combination of the state.

\begin{lem}\label{lem:state-vs-energy}
  Suppose the system behaves ideally (with no environment uncertainty and measurement error).
  The magnitude of two arbitrary states $\mathbf{q}_{k,0},\mathbf{q}_{k,1}$ described in~(\ref{eq:state-details}) along their evolution~(\ref{eq:state-unct}) at time instant $k$ depends on the instantaneous energy consumption
  \begin{equation}
    \|\mathbf{q}_{k,0}\|\geq\|\mathbf{q}_{k,1}\|\iff y_{k,0}\geq y_{k,1}.
  \end{equation}
\end{lem}
\begin{proof}
  *\\
\end{proof}

>>>>

The proposed algorithm alters the energy consumption dynamically by means of mission-specific parameters: the Quality of Service (QoS) of the onboard computations, and the trajectory-explicit equations (TEEs) of the path. Our goal is a mission extension by optimizing both QoS and TEEs as the UAV flies and its batteries drain. The algorithm optimizes QoS requiring the UAV to include robot operating system (ROS) nodes. It optimizes TEEs--an abstraction of the path to follow--and guides the UAV using a vector field that converges smoothly to the path. Physically, one can use a feature detection node and a set of ellipses, with the algorithm controlling the processing rate and the length of the semi-major and -minor axis. Data using a commercial fixed-wing UAV indicates a potential extension of up to 13 minutes over an hour by merely switching to the lowest QoS.

@@

The remainder of the paper is organized as follows. The dynamic mission planning problem is set in Section~\ref{sec:prob}, along a suitable model for the energy and position of the UAV. The algorithm that uses the model and solves the problem is proposed in Section~\ref{sec:algo}. Section~\ref{sec:experimental} presents a fixed-wing UAV flying in an agricultural scenario showcasing the performance. The paper finishes with some conclusions in Section~\ref{sec:conclusion}.

@@

The UAV flies in ellipses shifted in time--a survey pattern optimized for the craft's dynamics--detects obstacles using a convolutional neural network (CNN) and informs grounded mobile robots employed for future harvesting.

@@

We use the state to estimate the UAV's position in space and the energy in time. Starting with a position $\mathbf{p}\in\mathbb{R}^3$ in the 3D Euclidean space w.r.t. some inertial navigation frame $\mathcal{O}_W$, we build a guidance action--which allows the \emph{position evolution} along the 3-axis--to the path to follow using a vector field~\cite{de2017guidance}. The path is a set of generic continuous twice differentiable TEEs $\varphi_k:\mathbb{R}^3\rightarrow\mathbb{R}$. To this end, the TEE is satisfied $\varphi(\mathbf{p})\rightarrow 0$ for all the approaching points.

We build the \emph{energy evolution} $\mathbf{q}\in\mathbb{R}^j$ using Fourier analysis (the meaning of $j$ is clarified to the reader in Subsection~\ref{sec:energy-model}) and decompose such evolution in the energy due to the trajectory (depending on TEEs), and computations (on QoS)--an approach adapted from our earlier work on computational energy analysis~\cite{seewald2019coarse, seewald2019component}, and energy estimation of a fixed-wing UAV~\cite{seewald2020mechanical}.

\subsection{Position evolution and guidance action}
\label{sec:position}

We consider a non-holonomic 2D model of the UAV flying at an assigned altitude $h\in\mathbb{R}_{>0}$ (the requirement is eased in Section~\ref{sec:algo} to a set of altitudes $\mathbf{h}$)
\begin{equation}\label{eq:uav-kinem}\begin{cases}
  \dot{\mathbf{p}}(t)&=s\Psi(\psi(t))+\mathbf{d}(t)\\
  \dot{\psi}(t)&=u(\mathbf{p}(t))
\end{cases},
\end{equation}
where $\mathbf{p}\in\mathbb{R}^2$, $s\in\mathbb{R}$ is the airspeed (assumed constant), $\psi\in(-\pi,\pi]$ the attitude yaw angle and $\Psi(\psi)=\begin{bmatrix}\cos{\psi} & \sin{\psi}\end{bmatrix}^T$, $\mathbf{d}\in\mathbb{R}^2$ is the wind vector (assumed $\|\mathbf{d}\|<s$, i.e., the constant airspeed is greater than the norm of the wind), and $u\in\mathbb{R}$ the guidance action--the angular velocity $\dot{\psi}$.

Let us define $\mathcal{P}_k$, the area that discloses all the possible deviations form the $k$-th TEE 
\begin{equation}\label{eq:area}
  \mathcal{P}_k:=\{\mathbf{p}:\underline{c}_k\leq\varphi_k(\mathbf{p})\leq\overline{c}_k\},
\end{equation}
where $\underline{c}_k\in\mathbb{R}_{\leq 0},\overline{c}_k\in\mathbb{R}_{\geq 0}$ are the constraints: $\underline{a},\overline{a}$ returns the upper and lower bound of $a$ from the user-defined mission specification, a lookup table.

The algorithm uses the concept to select $c$ with the highest energy value under the energy budget constraints. Assume at time $k$ the algorithm selects the TEE $\varphi_k$ with the highest energy value being $c_k$, we design a guidance action $u$ that heads to such TEE by minimizing the norm $\|\varphi_k(\mathbf{p})-c_k\|$.

Let us define $\varPhi:=\varphi_k(\mathbf{p})-c_k$. The direction to the TEE to follow can be expressed as the desired velocity vector
\begin{equation}\label{eq:pd}
  \dot{\mathbf{p}}_d(\mathbf{p}):=E\nabla\varPhi-k_e\varPhi\nabla\varPhi,\,\,\,E=\begin{bmatrix}
    0&1\\-1&0
  \end{bmatrix},
\end{equation}
where $\nabla\varPhi\in\mathbb{R}^2$ is defined as the gradient of $\varphi$ at the point $\mathbf{p}$ (i.e., its vector field), $E$ specifies the tracking direction, and $k_e\in\mathbb{R}_{\geq 0}$ the gain to adjusts the speed of convergence.

@@

The energy evolution analysis necessitates the following realistic assumption.
\begin{assm}\label{assm:periodic} The position evolution is periodic, in the sense that it presents repetitive patterns.
\end{assm}
Formally, given two time instants $k_1,k_2\in\mathbb{R}_{\geq 0}$ such that $k_1>k_2$ and a value $n\in\mathbb{R}_{\geq 0}$, there exist an arbitrary constant displacement $e\in\mathbb{R}$ such that
\begin{equation}
  \|\mathbf{p}(t)-\mathbf{p}(t+n)\|= e\,\,\,\forall t\in\{k_1,...,k_2\}.
\end{equation}
We show in Section~\ref{sec:experimental} the assumption being eased in practice to a set $\mathbb{E}\subset\mathbb{R}$.

Let us consider a Fourier series $f:\mathbb{R}_{\geq 0}\rightarrow\mathbb{R}$ of an arbitrary order $r\in\mathbb{Z}_{\geq 0}$
\begin{equation}\label{eq:fourier}
  f(t)=\sum_{n=0}^{r}{a_n\cos{\frac{nt}{\xi}}+b_n\sin{\frac{nt}{\xi}}},
\end{equation}
where $\xi\in\mathbb{R}$ is the characteristic time, and $a_n, b_n\in\mathbb{R}$ for $n\in\{0,\dotsc,r\}$ the Fourier series coefficients.

The non-linear model in Equation~(\ref{eq:fourier}) can be expressed using an equivalent linear time-invariant state-space model
\begin{equation}\label{eq:state-perf}\begin{cases}
  \dot{\mathbf{q}}(t)&=A\mathbf{q}(t)+B\mathbf{u}(t)\\%+w(t)\\
  y(t)&=C\mathbf{q}(t)%+v(t)
\end{cases},\end{equation}
where $y(t)\in\mathbb{R}_{\geq 0}$ is the instantaneous energy consumption of the system being controlled. We prove formally in the Theorem~\ref{thm:state-vs-energy} the instantaneous energy being obtained as a linear combination of the state. The control $\mathbf{u}$ along with the input matrix $B$ are defined later in Subsection~\ref{sec:control-action}, the state $\mathbf{q}$ mimics the original Fourier series coefficients, and

\begin{equation}\label{eq:state-details}\begin{split}
  \mathbf{q}(t)&=\left[\begin{array}{cccccc}
    \alpha_0 & \alpha_1 & \beta_1 & \cdots & \alpha_r & \beta_r
  \end{array}\right]^T,\\
  A&=\left[\begin{array}{cccc}
    1&    &       &  \\
     & A_1&       &  \\
     &    & \ddots&  \\
     &    &       & A_r 
  \end{array}\right],\,A_n=\begin{bmatrix}0 & \frac{n}{\xi} \\ -\frac{n^2}{\xi^2} & 0\end{bmatrix},\\
  C&=\left[\begin{array}{cccccc}
    1 & 1 & 0 & \cdots & 1 & 0
  \end{array}\right],
\end{split}\end{equation}
where $\mathbf{q}(t)\in\mathbb{R}^j$ given $j:=2r+1$, $A\in\mathbb{R}^{j\times j}$ is the state transmission matrix, and $C\in\mathbb{R}^j$ is the output matrix. In matrix $A$, the first value is one, $A_n$ farther on the diagonal, and zero for the remaining.

We further model the system of Equation~(\ref{eq:state-perf}) plus disturbances with the following discrete time model
\begin{equation}\label{eq:state-unct}
  \begin{cases}
  \mathbf{q}_{k+1}&=A\mathbf{q}_{k}+B\mathbf{u}_{k}+\mathbf{w}_k\\
  y_k&=C\mathbf{q}_k+v_k
  \end{cases},
\end{equation}
where $\mathbf{w}_k\in\mathbb{R}^j,v_k\in\mathbb{R}$ account for unknown state and output disturbances. In the case of the system behaving ideally (i.e., with zero disturbances), we expect a set of states evolving accordingly to their output. Such observation is summarized in the following Lemma.

\begin{lem}\label{lem:state-vs-energy}
  Suppose the system of Equation~(\ref{eq:state-unct}) evolves with no disturbances ($\mathbf{w}=\mathbf{0},v=0$) and Assumption~\ref{assm:periodic} holds, then
  \begin{equation}
    \|\mathbf{q}_{k,0}\|\geq\|\mathbf{q}_{k,1}\|\iff y_{k,0}\geq y_{k,1}.
  \end{equation}
\end{lem}
\begin{proof}
  \begin{center}\vspace{.2cm}``\emph{The easy proof is trivial and is left as an exercise to the reader :P}''\vspace{.6cm}\end{center}
\end{proof}

@@
